Estimating Learning Dynamics in the Routing Game

### 2015-10-14 paper/paper.tex
Updates to sections 1, 2, 3 and bibliography

\documentclass{sig-alternate-ipsn13}
\usepackage{ntheorem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{todonotes}

@@

\maketitle
\begin{abstract}
  The routing game models congestion on transportation and communication networks.
  We consider an online learning model of player dynamics: at each iteration, every player chooses a route (or a probability distribution over routes), then the joint decision of all players determines the costs of each path, which are then revealed to the players. We first review convergence guarantees of such online learning dynamics. Then, we consider the following estimation problem: given a sequence of player decisions and the corresponding costs, we would like to fit the learning model parameters to these observations. We consider in particular entropic mirror descent dynamics, and develop a numerical solution to the estimation problem.

  We demonstrate this method using data collected from a routing game experiment: we develop a web interface to simulate the routing game. When players log in to the interface, they are assigned an origin and destination on the graph. They can choose, at each iteration, a distribution over their available routes, and each player seeks to minimize her own cost. We collect a data set using this interface, then we apply the proposed method to fit the learning model parameters.
  We observe in particular that after an exploration phase, the joint decision of the players remains within a small distance of the Nash equilibrium. We also use the estimated model parameters to predict the flow distribution over routes, and compare these predictions to the actual distribution.
  Finally, we discuss some of the qualitative implications of the experiment, and give directions for future research.

@@

%============================================================================================
\section{Introduction}
The routing game is a non-cooperative game that models congestion in transportation and communication networks. The game is given by a directed graph that represents the network, and each player is given by a source node and destination node, and seeks to send traffic (either packets in a communication setting, or cars in a  transportation setting) while minimizing the total delay of that traffic. The delay is determined by the joint decision of all players, such that whenever an edge has high load, it becomes congested and any traffic using that edge incurs additional delay.

This model of congestion is simple yet powerful, and routing games have been studied extensively since the seminal work of Beckman~\cite{beckmann1955studies}. The Nash equilibria of the game are simple to characterize, and have been used to quantify the inefficiency of the network, using the price of anarchy~\cite{roughgarden2002bad}. However, the Nash equilibrium concept may not offer a good descriptive model of actual behavior of players, as argued by many authors, including for example~\cite{fox2013population}. Besides the assumption of rationality, which can be questioned, the Nash equilibrium assumes that players have a complete description of the structure of the game, their own cost functions, and those of other players. This model is arguably not very realistic for the routing game, as one does not expect users of a network to have an accurate representation of the cost function on every edge of the network, or of the other users of the network.
One alternative to the Nash equilibrium concept is a model of repeated play, sometimes called learning models or adjustment models. In such models, one assumes that each player makes decisions iteratively, and uses the outcome of each iteration to adjust their next decision. Formally, if $x_k^{(t)}$ is the decision of player $k$ at iteration $t$, and $\ell^{(t)}_k$ is the corresponding vector of costs, then player $k$ faces a sequential decision problem in which she chooses $x^{(t)}_k$ then observes $\ell_k^{(t)}$. These sequential decision problems are coupled through the cost functions, since $\ell_k^{(t)}$ depends not only on $x_k^{(t)}$ but also on $x_{k'}^{(t)}$ for $k' \neq k$ (but players do not necessarily model this coupling). Such models have a long history in game theory, and date back to the work of Hannan~\cite{hannan1957approximations} and Blackwell~\cite{blackwell1956analog}. In recent years, there has been a resurgence of research on the topic of learning in games using sequential decision problems, see for example~\cite{cesa2006prediction} and the references therein.

When designing a model of player decisions, many properties are desirable. Perhaps the most important property is that the dynamics should be consistent with the equilibrium of the game, in the following sense: asymptotically, one should expect the learning dynamics to converge to the equilibrium of the full information, one-shot game (be it Nash equilibrium or other, more general equilibrium concepts). In this sense, players ``learn'' the equilibrium asymptotically. Much progress has been made in recent years in characterizing classes of learning dynamics which are guaranteed to converge to an equilibrium set~\cite{freund1999adaptive, hart2001general, hart2005adaptive, fox2013population}. In particular for the routing game, different models of learning have been studied for example in~\cite{fischer2004evolution,blum2006routing,kleinberg2009multiplicative,krichene2015learning,krichene2015SMD}.

In this paper, we briefly review some of the known models of learning in the routing game. We focus in particular on the mirror descent model used in~\cite{krichene2015MD}. This model describes the learning dynamics as solving, at each step, a simple minimization problem parameterized by a learning rate $\eta$. Formally, the decision at iteration $t+1$ is obtained by solving
\[
x^{(t+1)}_k(\eta^{(t)}_k) = \argmin_{x_k \in \Delta^{\Acal_k}} \eta^{(t)}_k\braket{\ell^{(t)}_k}{x} + D_{\psi_k}(x_k, x^{(t)}_k),
\]
where $\psi_k$ is a distance generating function with corresponding Bregman divergence $D_{\psi_k}$, and $\eta^{(t)}_k$ is a learning rate. The model is reviewed in detail in Section~\ref{sec:model}. Intuitively, minimizing the first term $\braket{\ell^{(t)}_k}{x}$ will assign traffic to the routes which currently have minimal cost, and minimizing the second term $D_{\psi_k}(x_k, x^{(t)}_k)$ will keep the traffic assignment at its current value. Minimizing the linear combination trades-off both terms, and the learning rate $\eta_k^{(t)}$ determines how aggressive the player is when updating her strategy: a small learning rate results in a small change in strategy (i.e. $x^{(t+1)}_k$ is close to $x_k^{(t)}$), while a large learning rate results in a significant change.

Motivated by this interpretation of learning rates, we propose in Section~\ref{sec:estimation}, the following estimation problem: given a sequence of player decisions $(x^{(t)}_k)$, and the sequence of corresponding losses $(\ell_k^{(t)})$, can we fit a learning model to these observations? A simple approach is to assume that the player is using a given distance generating function $\psi_k$, and estimate $\eta_k$ for example by minimizing the distance between the observed decision $\bar x^{(t+1)}_k$, and the decision predicted by the model, $x^{(t+1)}_k(\eta_k)$. More precisely, we can choose $\eta^{(t)}_k$ to minimize $D_{\psi_k}(\bar x^{(t+1)}_k, x^{(t+1)}_k(\eta))$. We show that in the entropic case (when $\psi_k$ is the negative entropy), this problem is convex, thus $\eta_k^{(t)}$ can be estimated efficiently e.g. by using gradient descent. This method allows us to estimate one parameter $\eta_k^{(t)}$ per iteration $t$. When we have a sequence of observations available, it can be desirable to control the complexity of the model by assuming a parameterized sequence of learning rates, instead of estimating each term separately. Thus, we propose a second method which assumes that the learning rate is of the form $\eta^{(t)}_k = \eta^{(0)}_k t^{-\alpha_k}$, with $\alpha_k \in (0, 1)$. The resulting estimation problem is non-convex in general, but since it is a two dimensional problem, it can be solved efficiently. Finally, we consider a family of distance generating functions $\psi_\epsilon$, parameterized by $\epsilon$, that can be viewed as a generalization of the negative entropy function. These generalized entropy functions also offer some desirable properties that we discuss in Section~\ref{sec:estimation}. We also briefly discuss potential uses for the estimated model: for example, the model can be used simply to predict the decision of the players over the next few iterations, by propagating the model forward with the estimated values of the parameters; more generally, the model can be used to formulate a receding-horizon optimal control problem, by using the current estimate of the model as a plant in the control problem.

In the second part of the paper, we present an experimental setting which we developed to collect data on routing decisions. We developed a web interface in which a master user can create an instance of the routing game by defining a graph and cost functions on edges of the graph. Then other users can connect to the interface as players. The game then proceeds similarly to our learning model: at each iteration, every player chooses a flow distribution on their available routes (using a graphical user interface with sliders), then their decisions are sent to a backend server, which computes the total cost of each route, and sends back this information to each player. In Section~\ref{sec:experiment}, we describe the experimental setting, some implementation details, as well as the nature of the collected data. We then use this data to run the estimation tasks which were proposed in Section~\ref{sec:estimation}, and give some qualitative and quantitative insights into the behavior of players. In particular, we observed that in the first few iterations, the flow distributions oscillate, which corresponds to a high value of estimated learning rates. For later iterations, the flow distributions are, in general, close to equilibrium, and the learning rates are lower, although some players may occasionally move the system away from equilibrium by performing an aggressive update (corresponding to a high learning rate). It is also interesting to observe that in some rare cases, the estimate of the learning rate is negative, which means that the player updated her strategy by assigning more traffic to routes with higher cost, a counter-intuitive behavior which is hard to model. We also solve the decay rate estimation problem and compare the decay rates of the different players, then use the estimated parameters to predict the traffic assignment over the next few iterations, and comment on the quality of the prediction. We conclude in Section~\ref{sec:conclusion} by summarizing our results and giving directions for future research.

@@

\subsection{The routing game}
\label{sec:routing_game}
The routing game is played on a directed graph $\Gcal = (V, E)$, where $V$ is a vertex set and $E \subset V \times V$ is an edge set. The players will be indexed by $k \in \{1, \dots, K\}$, where every player is given by an origin vertex $o_k \in V$, a destination vertex $d_k \in V$, and a traffic mass $m_k \geq 0$ that represents the total traffic that the player needs to send from $o_k$ to $d_k$. The set of available paths connecting $o_k$ to $d_k$ will be denoted by $\Pcal_k$, and the action set of player $k$ is simply the probability simplex over $\Pcal_k$, which we denote by $\Delta_k = \{x \in \Rbb_+^{\Pcal_k} : \sum_{p \in \Pcal_k} x_p = 1\}$. In other words, each player chooses a distribution over their available paths, and their traffic is allocated to paths according to that distribution. We will denote by $x_k\in \Delta^{\Pcal_k}$ the distribution of player $k$. Note that $x_k$ is a distribution vector, so the vector of actual flows is the scaled vector $m_k x_k$. The joint decision of all players is denoted by $x = (x_1, \dots, x_K)$. The costs of the players are then determined as follows:
\begin{itemize}
\item The cost on an edge $e$ is $c_e(\phi_e(x))$, where $c_e(\cdot)$ is a given, increasing function, and $\phi_e(x)$ is the total traffic flow on edge $e$ induced by the distribution $x$, obtained simply by summing all the path flows that go through that edge, i.e. $\phi_e(x) = \sum_k \sum_{p \in \Pcal_k} m_k x_{k, p}$.
\item The cost on a path $p \in \Pcal_k$ is denoted by $\ell_{k, p}(x)$, and is the sum of edge costs along the path, i.e. $\ell_{k, p}(x) = \sum_{e \in p} c_e(\phi_e(x))$.
\item The cost for player $k$ is the total path cost for all the traffic sent by player $k$, i.e. $\sum_{p \in \Pcal_k} m_k x_{k, p} \ell_{k, p}(x)$. This is simply the inner product between the flow vector $m_k x_k$ and the path loss vector $\ell_k(x)$, which we denote by $\braket{\ell_k(x)}{x_k}$.
\end{itemize}
\begin{remark}[A note on the player model] Some formulations of the routing game, e.g.~\cite{sandholm2001potential,krichene2015learning}, formulate the game in terms of populations of players, such that each population is an infinite set of players with the same origin and destination. This assumes that each player contributes an infinitesimal amount of flow, so each player can pick a single path. In our model, each player is macroscopic, and can split its traffic across multiple routes. Both models are equivalent in terms of analysis, the only difference is in terms of interpretation of the model. We choose the finite player interpretation because it is more consistent with the experimental section of the paper, where we run the game with finitely many players.
\end{remark}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{definition}[Nash equilibrium]
A distribution $x^\star = (x_1^\star, \dots, x_K^\star)$ is a Nash equilibrium if it satisfies the following condition: for all other feasible  distributions $x = (x_1, \dots, x_K)$ and for all~$k$,
\[
\braket{\ell_k(x^\star)}{x_k - x_k^\star} \geq 0.
\]
\end{definition}
In words, $x^\star$ is a Nash equilibrium if for every player $k$, the expected cost under $x_k^\star$ is lower than the expected cost under any other distribution $x_k$. If we define the inner product $\braket{x}{\ell} = \sum_{k} \braket{x_k}{\ell_k}$, then this is equivalent to: $x^\star$ is an equilibrium if and only if $\braket{\ell(x^\star)}{x - x^\star} \geq 0$ for all feasible $x$. This variational inequality is, in fact, equivalent to the first-order optimality condition of the following potential function, usually referred to as the Rosenthal potential, in reference to~\cite{rosenthal1973class}:
\begin{proposition}[Existence of a convex potential]
Consider a routing game and define the following function
\[
f(x) = \sum_{e \in E} \int_0^{\phi_e(x)} c_e(u)du.
\]
Then $f$ is convex its gradient is $\nabla f(x) = \ell(x)$.
\end{proposition}
This result can be found for example in~\cite{roughgarden2002bad}. Due to the fact that the loss function $\ell(\cdot)$ coincides with the gradient field $\nabla f(\cdot)$ of the Rosenthal potential, the Nash condition can be rewritten as
\[
\braket{\nabla f(x^\star)}{x - x^\star} \geq 0 \ \forall \text{ feasible }x,
\]
and since $f$ is convex, this is a necessary and sufficient condition for optimality of $x^\star$ (see e.g. Section 4.2.3 in~\cite{boyd2010convex}). Therefore the set of Nash equilibria is exactly the set of minimizers of the convex potential~$f$. This is important both for computation (computing a Nash equilibrium can be done by minimizing a convex function), and for modeling: one can model player dynamics as performing a distributed optimization of the potential function. More precisely, if we adopt the point of view presented in the introduction, wherein each player faces a sequential decision problem, and plays $x^{(t)}_k$ then observes $\ell_k(x^{(t)})$, then this corresponds to a first-order distributed optimization of the function~$f$, where each player is responsible for updating the variables $x_k^{(t)}$, and observes, at each iteration, the partial gradient $\ell_k(x^{(t)}) = \nabla_{x_k} f(x^{(t)})$. Using this connection to distributed optimization, a model of player dynamics was proposed in~\cite{krichene2015MD}. We review the model in the next Section.

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{The learning model: mirror descent dynamics}
\label{sec:learning}
We will consider the model of distributed learning proposed in~\cite{krichene2015MD}. Each player is assumed to perform a mirror descent update given by the following algorithm:
\begin{algorithm}[H]
\begin{algorithmic}
\FOR{each $t \in \{1, 2, \dots \}$}
\STATE Play $x^{(t)}_k$,
\STATE Observe $\ell_k^{(t)} = \nabla_{x_k} f(x^{(t)})$,
\STATE Update
\begin{equation}
x^{(t+1)}_k = \argmin_{x_k \in \Delta^{\Pcal_k}} \eta_k^{(t)} \braket{\ell_k(x^{(t)})}{x_k} + D_{\psi_k}(x_k, x^{(t)}_k)
\label{eq:MD_update}
\end{equation}
\ENDFOR
\end{algorithmic}
\caption{Distributed mirror descent dynamics with DGF $\psi_k$ and learning rates $\eta_k^{(t)}$.}
\label{alg:MD}
\end{algorithm}
In the update equation~\eqref{eq:MD_update}, $D_{\psi_k}(x, x^{(t)}_k)$ is the Bregman divergence between the distributions $x_k$ and $x_k^{(t)}$, defined as follows: $D_\psi(x, y) = \psi(x) - \psi(y) - \braket{\nabla \psi(y)}{x - y}$, for a strongly convex function $\psi$, called the distance generating function (DGF). Some special cases include
\begin{itemize}
\item The Euclidean case: if $\psi(x) = \frac{1}{2}\|x\|^2_2$, then $D_\psi(x, y) = \frac{1}{2}\|x - y\|_2^2$. In this case, mirror descent reduces to the projected gradient descent algorithm.
\item The entropic case: if $\psi(x) = - H(x)$ where $H(x) = -\sum_{p} x_p \ln x_p$ is the negative entropy, then $D_\psi(x, y) = \sum_p x_p \ln \frac{x_p}{y_p}$ is the Kullback-Leibler (KL) divergence from $x$ to $y$. In this case, the mirror descent algorithm is sometimes called the entropic descent~\cite{beck2003mirror}, or exponentiated gradient descent~\cite{kivinen1997exponentiated}.
\end{itemize}
The mirror descent method is a general method for convex optimization proposed in~\cite{nemirovski1983problem}. The model in Algorithm~\ref{alg:MD} is a distributed version of mirror descent, applied to the potential function $f$. To give some intuition of the method, the first term $\braket{\ell_k^{(t)}}{x_k}$ in the minimization problem~\eqref{eq:MD_update} can be thought of as a linear approximation of the potential function (since $\ell(x) = \nabla f(x)$), and the second term $D_\psi(x_k, x^{(t)}_k)$ penalizes deviations from the previous iterate. The learning rate $\eta_k^{(t)}$ determines the tradeoff between the two terms, and can be thought of as a generalized step size: a smaller $\eta_k$ results in a distribution which is closer to the current $x_k^{(t)}$. Thus, from the potential function point of view, the player minimizes a linearization of the potential plus a Bregman divergence term that keeps $x_k$ close to $x_k^{(t)}$. From the routing game point of view, the first term $\braket{\ell^{(t)}_k}{x_k}$ corresponds to putting weight on the paths that have smaller cost during the previous iteration, and the second term keeps the distribution at its current value.

>>>>

\documentclass{sig-alternate-ipsn13}
\usepackage{ntheorem}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}

@@

\maketitle
\begin{abstract}
  The routing game models congestion in transportation networks, communication networks, and other cyber physical systems in which agents compete for shared resources.
  We consider an online learning model of player dynamics: at each iteration, every player chooses a route (or a probability distribution over routes, which corresponds to a flow allocation over the physical network), then the joint decision of all players determines the costs of each path, which are then revealed to the players.

  We pose the following estimation problem: given a sequence of player decisions and the corresponding costs, we would like to estimate the learning model parameters. We consider in particular entropic mirror descent dynamics, reduce the estimation problem to estimating the learning rates of the players, and develop a numerical solution to the estimation problem.

  We demonstrate this method using data collected from a routing game experiment, played by human participants: We develop a web interface to implement the routing game. When players log in to the web application, they are assigned an origin and destination on the graph. They can choose, at each iteration, a distribution over their available routes, and each player seeks to minimize her own cost. We collect a data set using this interface, then apply the proposed method to estimate the learning model parameters.
  We observe in particular that after an exploration phase, the joint decision of the players remains within a small distance of the Nash equilibrium. We also use the estimated model parameters to predict the flow distribution over routes, and compare these predictions to the actual distribution.
  Finally, we discuss some of the qualitative implications of the experiments, and give directions for future research.

@@

%============================================================================================
\section{Introduction}
The routing game is a non-cooperative game that models congestion in many cyber physical systems (CPS) in which non-cooperative agents compete for shared resources, such as transportation networks (the resources being roads) and communication networks (the resources being communication links)~\cite{beckmann1955studies,roughgarden2007,ozdaglar2007incentives}. The game is played on a directed graph that represents the network, and each player is given by a source node and destination node, and seeks to send traffic (either packets in a communication setting, or cars in a  transportation setting) while minimizing the total delay of that traffic. The delay is determined by the joint decision of all players, such that whenever an edge has high load, it becomes congested and any traffic using that edge incurs additional delay, defined by a congestion function that models the underlying physical process. This model of congestion is simple yet powerful, and routing games have been studied extensively since the seminal work of Beckman~\cite{beckmann1955studies}.

\subsection{Learning models and convergence to Nash equilibria}
The Nash equilibria of the game are simple to characterize, and have been used to quantify the inefficiency of the network, using the price of anarchy~\cite{roughgarden2007}. However, the Nash equilibrium concept may not offer a good descriptive model of actual behavior of players. Besides the assumption of rationality, which can be questioned~\cite{simon1955behavioral}, the Nash equilibrium assumes that players have a complete description of the structure of the game, their own cost functions, and those of other players. This model is arguably not very realistic for the routing game, as one does not expect users of a network to have an accurate representation of the cost function on every edge of the network, or of the other users of the network.
One alternative to the Nash equilibrium concept as a descriptive model of players is a model of repeated play~\cite{marden2013game,fox2013population}, sometimes called learning models~\cite{cesa2006prediction} or adjustment models~\cite{fudenberg1998theory}. In such models, one assumes that each player makes decisions iteratively, and uses the outcome of each iteration to adjust their next decision. Formally, if $x_k^{(t)}$ is the decision of player $k$ at iteration $t$, and $\ell^{(t)}_k$ is the corresponding vector of costs (delays), then player $k$ faces a sequential decision problem in which she iteratively chooses $x^{(t)}_k$ then observes $\ell_k^{(t)}$. These sequential decision problems are coupled through the cost functions, since $\ell_k^{(t)}$ depends not only on $x_k^{(t)}$ but also on $x_{k'}^{(t)}$ for $k' \neq k$ (but players do not necessarily model this coupling). Such models have a long history in game theory, and date back to the work of Hannan~\cite{hannan1957approximations} and Blackwell~\cite{blackwell1956analog}. In recent years, there has been a resurgence of research on the topic of learning in games using sequential decision problems, see for example~\cite{cesa2006prediction} and references therein.

When designing a model of player decisions, many properties are desirable. Perhaps the most important property is that the dynamics should be consistent with the equilibrium of the game, in the following sense: Asymptotically, one should expect the learning dynamics to converge to the equilibrium of the full information, one-shot game (be it Nash equilibrium or other, more general equilibrium concepts). In this sense, players ``learn'' the equilibrium asymptotically. Much progress has been made in recent years in characterizing classes of learning dynamics which are guaranteed to converge to an equilibrium set~\cite{freund1999adaptive, hart2001general, hart2005adaptive, fox2013population}. In particular for the routing game, different models of learning have been studied for example in~\cite{fischer2004evolution,blum2006routing,kleinberg2009multiplicative,krichene2015learning,krichene2015SMD}, with different convergence guarantees.

\subsection{A mirror descent model of learning}
We will focus in particular on the mirror descent model used in~\cite{krichene2015MD}, since it offers a large family of models that have strong convergence guarantees to Nash equilibria. This model describes the learning dynamics as solving, at each step, a simple minimization problem parameterized by a learning rate $\eta$. Formally, the decision at iteration $t+1$ is obtained by solving
\[
x^{(t+1)}_k(\eta^{(t)}_k) = \argmin_{x_k \in \Delta^{\Acal_k}} \eta^{(t)}_k\braket{\ell^{(t)}_k}{x} + D_{\psi_k}(x_k, x^{(t)}_k),
\]
where $\psi_k$ is a distance generating function with corresponding Bregman divergence $D_{\psi_k}$, and $\eta^{(t)}_k$ is a learning rate. Intuitively, minimizing the first term $\braket{\ell^{(t)}_k}{x}$ will assign traffic to the routes which currently have minimal cost, and minimizing the second term $D_{\psi_k}(x_k, x^{(t)}_k)$ will keep the traffic at its current value. Minimizing the linear combination trades-off both terms, and the learning rate $\eta_k^{(t)}$ determines how aggressive the player is when updating her strategy: A small learning rate results in a small change in strategy (i.e. $x^{(t+1)}_k$ is close to $x_k^{(t)}$), while a large learning rate results in a significant change.

\subsection{Estimating the learning rates}
Motivated by this interpretation of the learning dynamics, we propose the following estimation problem: Given a sequence of player decisions $(x^{(t)}_k)$, and the sequence of corresponding costs $(\ell_k^{(t)})$, can we estimate the learning model parameters to fit these observations? These quantities are effectively measured in our experimental setting using the routing interface, and can be measured on transportation networks using existing traffic monitoring and forecasting systems, such as the Mobile Millennium system~\cite{bayen2011mobile} or the Grenoble Traffic Lab~\cite{canudasdewit2015grenoble}.

Our proposed approach is to assume that the player is using a given distance generating function $\psi_k$, and estimate $\eta_k$ for example by minimizing the distance between the observed decision $\bar x^{(t+1)}_k$, and the decision predicted by the model, $x^{(t+1)}_k(\eta_k^{(t)})$. More precisely, we can choose $\eta^{(t)}_k$ to minimize $D_{\psi_k}(\bar x^{(t+1)}_k, x^{(t+1)}_k(\eta_k^{(t)}))$. We show that in the entropic case (when $\psi_k$ is the negative entropy), this problem is convex, thus $\eta_k^{(t)}$ can be estimated efficiently e.g. by using gradient descent. This method allows us to estimate one parameter $\eta_k^{(t)}$ per iteration $t$ and per player $k$. When we have a sequence of observations available, it can be desirable to control the complexity of the model by assuming a parameterized sequence of learning rates, instead of estimating each term separately. Thus, we propose a second method which assumes that the learning rate is of the form $\eta^{(t)}_k = \eta^{(0)}_k t^{-\alpha_k}$, with $\alpha_k \in (0, 1)$. The resulting estimation problem is non-convex in general, but since it is a two dimensional problem, it can be minimized efficiently. Finally, we consider a family of distance generating functions $\psi_\epsilon$, parameterized by $\epsilon$, that can be viewed as a generalization of the negative entropy function. These generalized entropy functions offer some desirable properties that will be discussed in more detail.

\subsection{Summary of contributions and organization of the article}
Our main contributions are to
\setlist[enumerate]{itemindent=\dimexpr\labelwidth+\labelsep\relax,leftmargin=0pt}
\begin{enumerate}
\item Pose the learning rate estimation problem, and show that it is convex problem in the entropic case. We also give an example application of the estimated model: It can be used to predict the decision of the players over the next few iterations, by propagating the model forward with the estimated values of the parameters.
\item Develop a routing game system in order to collect data on routing decisions. We developed a web interface in which a master user can create an instance of the routing game by defining a graph and cost functions on edges of the graph. Then other users can connect to the interface as players. The game then proceeds similarly to our learning model: At each iteration, every player chooses a flow distribution on their available routes (using a graphical user interface with sliders), then their decisions are sent to a backend server, which computes the total cost of each route, and sends this information back to each player.
\item Apply the proposed methods to the data collected from the routing game system, and give quantitative and qualitative insights into the decision dynamics of human players. In particular, we observed that in the first few iterations, the flow distributions oscillate, which corresponds to a high value of estimated learning rates. For later iterations, the flow distributions are, in general, close to equilibrium, and the learning rates are lower, although some players may occasionally move the system away from equilibrium by performing an aggressive update (high learning rate). It was also interesting to observe that in some rare cases, the estimate of the learning rate is negative, which means that the player updated her strategy by assigning more traffic to routes with higher cost, a counter-intuitive behavior which is hard to model. Finally, the prediction with the estimated model performs surprisingly well, especially for short term prediction (up to $4$ iterations), which indicates that the mirror descent model is a good descriptive and predictive model for player behavior in this setting.
\end{enumerate}

The remainder of the article is organized as follows: In Section~\ref{sec:model}, we formally define the routing game and review the characterization of its equilibrium set, then define the mirror descent dynamics and review its convergence guarantees. In Section~\ref{sec:estimation}, we pose the learning rate estimation problem in the entropy case, then extend it to the generalized entropy case. We also briefly discuss the traffic prediction problem. In Section~\ref{sec:experiment}, we describe the experimental setting, some implementation details, as well as the nature of the collected data. We then use this data to run the estimation and prediction tasks in Section~\ref{sec:results}, comment on the quality of the prediction, and give some qualitative and quantitative insights into the decision dynamics (distance to equilibrium, comparison of learning decay rates). We conclude in Section~\ref{sec:conclusion} by summarizing our results and giving directions for future research.

@@

\subsection{The routing game}
\label{sec:routing_game}
The routing game is played on a directed graph $\Gcal = (V, E)$, where $V$ is a vertex set and $E \subset V \times V$ is an edge set. The players will be indexed by $k \in \{1, \dots, K\}$, where every player is given by an origin vertex $o_k \in V$, a destination vertex $d_k \in V$, and a traffic mass $m_k \geq 0$ that represents the total traffic that the player needs to send from $o_k$ to $d_k$. The set of available paths connecting $o_k$ to $d_k$ will be denoted by $\Pcal_k$, and the action set of player $k$ is simply the probability simplex over $\Pcal_k$, which we denote by $\Delta^{\Pcal_k} = \{x \in \Rbb_+^{\Pcal_k} : \sum_{p \in \Pcal_k} x_p = 1\}$. In other words, each player chooses a distribution over their available paths, and their traffic is allocated to paths according to that distribution. We will denote by $x_k\in \Delta^{\Pcal_k}$ the distribution of player $k$. Note that $x_k$ is a distribution vector, so the vector of actual flows is the scaled vector $m_k x_k$. The joint decision of all players is denoted by $x = (x_1, \dots, x_K)$. The costs of the players are then determined as follows:
\begin{enumerate}[label=\alph*)]
\item The cost on an edge $e$ is $c_e(\phi_e(x))$, where $c_e(\cdot)$ is a given, increasing function (this models the actual cost due to the physical process, for example delay on a road segment due to accumulation of cars), and $\phi_e(x)$ is the total traffic flow on edge $e$ induced by the distribution $x$, obtained simply by summing all the path flows that go through that edge, i.e. $\phi_e(x) = \sum_k \sum_{p \in \Pcal_k} m_k x_{k, p}$.
\item The cost on a path $p \in \Pcal_k$ is denoted by $\ell_{k, p}(x)$, and is the sum of edge costs along the path, i.e. $\ell_{k, p}(x) = \sum_{e \in p} c_e(\phi_e(x))$.
\item The cost for player $k$ is the total path cost for all the traffic sent by player $k$, i.e. $\sum_{p \in \Pcal_k} m_k x_{k, p} \ell_{k, p}(x)$. This is simply the inner product between the flow vector $m_k x_k$ and the path delay vector $\ell_k(x)$, which we denote by $\braket{\ell_k(x)}{x_k}$.
\end{enumerate}
\begin{remark}[A note on the player model] Some formulations of the routing game, e.g.~\cite{sandholm2001potential,krichene2015learning}, define the game in terms of \emph{populations} of players, such that each population is an infinite set of players with the same origin and destination. This assumes that each player contributes an infinitesimal amount of flow, so each player can pick a single path. In our model, each player is macroscopic, and can split its traffic across multiple routes. Both models are equivalent in terms of analysis, the only difference is in terms of interpretation of the model. We choose the finite player interpretation because it is more consistent with the experimental section of the article, where we run the game with finitely many players.
\end{remark}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{definition}[Nash equilibrium]
A distribution $x^\star = (x_1^\star, \dots, x_K^\star)$ is a Nash equilibrium if it satisfies the following condition: For all other feasible  distributions $x = (x_1, \dots, x_K)$ and for all~$k$, $\braket{\ell_k(x^\star)}{x_k - x_k^\star} \geq 0$.
\end{definition}
In words, $x^\star$ is a Nash equilibrium if for every player $k$, the expected cost under $x_k^\star$ is lower than the expected cost under any other distribution $x_k$. If we define the inner product $\braket{x}{\ell} = \sum_{k} \braket{x_k}{\ell_k}$, then this is equivalent to: $x^\star$ is an equilibrium if and only if $\braket{\ell(x^\star)}{x - x^\star} \geq 0$ for all feasible $x$. This variational inequality is, in fact, equivalent to the first-order optimality condition of the following potential function, usually referred to as the Rosenthal potential, in reference to~\cite{rosenthal1973class}:
\begin{proposition}[Existence of a convex potential]
\label{prop:potential}
Consider a routing game and define the following function
$
f(x) = \sum_{e \in E} \int_0^{\phi_e(x)} c_e(u)du.
$
Then $f$ is convex its gradient is $\nabla f(x) = \ell(x)$.
\end{proposition}
This result can be found for example in~\cite{roughgarden2007}. Due to the fact that the delay function $\ell(\cdot)$ coincides with the gradient field $\nabla f(\cdot)$ of the Rosenthal potential, the Nash condition can be rewritten as
$\braket{\nabla f(x^\star)}{x - x^\star} \geq 0$ for all feasible $x$,
and since $f$ is convex, this is a necessary and sufficient condition for optimality of $x^\star$ (see e.g. Section 4.2.3 in~\cite{boyd2010convex}). Therefore the set of Nash equilibria is exactly the set of minimizers of the convex potential~$f$. This is important both for computation (computing a Nash equilibrium can be done by minimizing a convex function), and for modeling: One can model player dynamics as performing a distributed optimization of the potential function. More precisely, if we adopt the point of view presented in the introduction, in which each player faces a sequential decision problem, and plays $x^{(t)}_k$ then observes $\ell_k(x^{(t)})$, then this corresponds to a first-order distributed optimization of the function~$f$, where each player is responsible for updating the variables $x_k^{(t)}$, and observes, at each iteration, the partial gradient $\ell_k(x^{(t)}) = \nabla_{x_k} f(x^{(t)})$. Using this connection to distributed optimization, a model of player dynamics was proposed in~\cite{krichene2015MD}. We review the model in the next Section.
